---
title: "Use BirdynamicPkg package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use BirdynamicPkg package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```


# Introduction
The R package 'BirdynamicPkg' enables modelling the demographic impact of windfarms on seabird species following the methodology developed in the Bird Dynamic project. Check out the 'Read me' file to know more.


# Download and load the package

```{r setup, echo=F}
#devtools::install_github("BaptisteAverly/BirdynamicPkg", ref="develop")
library(BirdynamicPkg)

library(stochLAB)
library(rhandsontable)
library(popbio)
library(plyr)
library(tidyverse)
library(plotly)
library(rmarkdown)
library(markdown)
library(DT)
library(leaflet) # Needed for interactive map
library(leafem) # Needed for addMouseCoordinates
library(gdistance) # Needed to calculate shortest path distance
library(sf) # Needed for GIS work
library(raster) # Needed for GIS work (extent function)
library(units)
library(htmlwidgets) # Needed for onRender in leaflet map
library(reshape2) # Needed to prepare output tables
library(nimble)

```


# Prepare data

INCLURE RECAP DES DONNEES A UTILISER

## Count data
TO COME (GISOM for now)

```{r count_data}
countData <-  read.csv2("../../Birdynamic/0.BV_Data_input/BD_Effectifs_clean.csv") %>% subset(., seafront=="atlantique")
colonies_all <- bdy_check_colonies(countData)
countData$colony_code <- colonies_all$colony_code[match(countData$colony,colonies_all$colony)]
```


## Species parameters
To run the analyses, parameters of the species of interest have to be included in the following tables. For each of these tables, the default values that we use in the Bird Dynamic Shiny app are included in the package, but you can decide to edit them with different values or adding species.

The table `vital_rates` includes the survival and reproduction parameters:
```{r data_vital}
vital_rates <- bdydata_vital_rates
head(vital_rates)
```


The table `foraging_ranges` includes the foraging range of each species:
```{r data_foraging}
foraging_ranges <- bdydata_foraging_ranges
head(foraging_ranges)
```


The table `seasons` includes the presence of each species at each month (see `help(bdydata_seasons)` to see details on attributes):
```{r data_seasons}
seasons <- bdydata_seasons
head(seasons)
```


## Mortality data
Mortality can be either calculated using the `stochLAB` package or directly loaded if they were calculated previously.

### Calculate mortality data
To create them with the `stoch_crm()` function from the `stochLAB` package, you must define the density of your species for each parc and month, and all the parc and seabird parameters. The values entered below are only for the example, do not consider them as verified default values. 

```{r calculate_mortality}

dens_dt <- data.frame(
  row.names = month.name,
  month=month.name,
  mean=c(5,5,5,10,10,10,10,10,10,5,5,5),
  sd=1
)

chord_profile <- data.frame(
  pp_radius=seq(0, 1, by=0.05),
  chord=c(0.69, 0.73, 0.79, 0.88, 0.96, 1, 0.98, 0.92, 0.85, 0.8, 0.75, 0.7, 0.64, 0.58, 0.52, 0.47, 0.41, 0.37, 0.3, 0.24, 0)
)

n_iterations <- 1000

morta_results <- stoch_crm(
  model_options = "1",
  n_iter = n_iterations,
  flt_speed_pars = data.frame("mean"=0.1, "sd"=0),
  body_lt_pars = data.frame("mean"=0.1, "sd"=0),
  wing_span_pars = data.frame("mean"=0.1, "sd"=0),
  avoid_bsc_pars = data.frame("mean"=0.99, "sd"=0),
  avoid_ext_pars = NULL,
  noct_act_pars = data.frame("mean"=0.2, "sd"=0),
  prop_crh_pars = data.frame("mean"=0,"sd"=0),
  bird_dens_opt = "tnorm",
  bird_dens_dt = dens_dt,
  flight_type = "flapping",
  prop_upwind = 0.5,
  n_blades = 3,
  air_gap_pars = data.frame("mean"=30,"sd"=0),
  rtr_radius_pars = data.frame("mean"=80, "sd"=0),
  bld_width_pars = data.frame("mean"=5.5,"sd"=0),
  bld_chord_prf = chord_profile,
  rtn_pitch_opt = "probDist",
  bld_pitch_pars = data.frame("mean"=2, "sd"=0.1),
  rtn_speed_pars = data.frame("mean"=10, "sd"=0.5),
  trb_wind_avbl = cbind(data.frame("month"=month.name),"pctg"=rep(95, 12)),
  trb_downtime_pars = cbind(data.frame("month"=month.name), "mean"=rep(6, 12), "sd"=rep(2, 12)),
  wf_n_trbs = 50,
  wf_width = 5,
  wf_latitude = 48.67,
  tidal_offset = 2,
  lrg_arr_corr = T,
  out_format = "draws",
  out_sampled_pars = FALSE,
  out_period = "months",
  season_specs = NULL,
  verbose = TRUE,
  log_file = NULL,
  seed = NULL
)

morta_format <- data.frame("espece_latin"="Larus fuscus",
                           "parc"="parc_1",
                           "mois"=rep(1:12, n_iterations),
                           "iteration"=rep(1:n_iterations, each=12),
                           "coefficient"=as.vector(t(morta_results[["opt1"]]))
                           )

head(morta_format)
```


### Load mortality data
Alternatively, mortality data that were compiled previously can be directly loaded, as long as they are formatted as in `bdydata_mortality_example` (which we will use in this vignette):

```{r load_mortality}
FormattedMortality <- bdy_check_mortality(bdydata_mortality_example, bdydata_mortality_example$espece_latin)$table
head(FormattedMortality)
```



## Parc data
Parc data must be provided to run the analyses, as a sf object (see `st_read(st_read)` for more details) using either points or polygons, projected in Lambert93 (EPSG:2154). As an example, we here use the `bdydata_parc_example` data included in the package and we apply the `bdy_clean_names()` function to the NAME column (to remove spaces and capital and ensure it matches with names in mortality data). They must include a column indicating the seafront (for France, it can be 'atlantique' or 'mediterranee'); this also has to be added as a column to the colony file.

```{r parc_example}
parcs_L93 <- bdydata_parcs_example %>% st_transform(., st_crs(2154))
parcs_L93$NAME <- bdy_clean_names(parcs_L93$NAME)
parcs_L93$seafront <- "atlantique"
parcs_L93

colonies_all$seafront <- "atlantique"
```


ADD Check mortality names


## Prepare coast line
A map of coastline is needed to create a cost matrix to calculate distance by sea, and to calculate the proportion of sea area around each colonies. We use data stored in the package and crop based on the location of colonies.

```{r countryShape}
crop_extent <- colonies_all %>% st_bbox() %>% st_as_sfc() %>% st_buffer(., 100000) %>% st_transform(., 4326)

countryShape <- bdydata_world_map %>%
  st_crop(., crop_extent) %>%
  st_transform(., st_crs(colonies_all))

ggplot()+
  geom_sf(data=countryShape)+
  geom_sf(data=colonies_all, col="darkred")+
  theme_minimal()

```



# Detailed analysis for one species

## Calculate distance between colonies and parcs
We calculate two types of distance: euclidian distance is used for species that cross land (e.g., gulls), shortest sea path is used for purely marine species (e.g., Northern Gannet). Both distances are calculated for all colonies (if doShpa is set to TRUE) with the `bdy_get_distances()` function, the type of distance to use then depends on species later in the analysis. Before calculating distances, a cost matrix based on coastline must be calculated with the `bdy_get_cost_raster()`.

```{r calc_dist}
cost_matrix <- bdy_get_cost_raster(countryShape, colonies_all)

all_distances <- bdy_get_distances(colonies=colonies_all,
                                   parcs = parcs_L93,
                                   costMatrix=cost_matrix,
                                   doShpa = T)

head(all_distances$eucl_dist)
head(all_distances$shpa_dist)
```


## Process count data
The next steps of the analysis are applied species by species. In this vignette, we show a detailed example for one species, followed by how you can loop the analysis for multiple species.

For this example, we will select the Northern Gannett (**Morus bassanus**) and process count data. The processing step includes an important step, which is to cluster close colonies (reported in the column `group`).
```{r select_species}
sp <- "Morus bassanus"

count_processed <- bdy_process_count_data(sp=sp,
                                          effec00=countData,
                                          colonies00=colonies_all,
                                          first_year=2009,
                                          last_year = 2021
)
```


## Apportionning and processing mortality data
The first step to apply is the apportionning, where mortalities due to collisions with wind turbines (modelled or loaded earlier in the pipeline) are attributed to the different colonies depending on their distance to the windfarms or the configuration of the coastline.

We first set the distance between parcs and colonies that include the selected species, keeping the shortest sea path if the species is purely marine, the euclidian distance otherwise.

```{r select_dist}
if(foraging_ranges$terrestrial_habits[foraging_ranges$species_latin==sp]){
  distances <- all_distances$eucl_dist
}else{
  distances <- all_distances$shpa_dist
}
distances <- distances[rownames(distances) %in% count_processed$colonies_sp$colony_code,]
```


We can then calculate the proportion of sea area around each colony, which is used in the apportionning step to account for the higher density of individuals at sea if the proportion of sea area around a colony is low (as the species is not using the terrestrial habitat). 

```{r sea_area}
sea_area <- bdy_calculate_sea_area(country_polygon=countryShape, 
                                   max_foraging_range = foraging_ranges$max_km[which(foraging_ranges$species_latin==sp)],
                                   colonies=count_processed$colonies_sp
                                   )

head(sea_area)
```


We can now apply the [bdy_apportionning()] function to run the apportionning calculations to set how mortality data should be distributed among colonies. The output includes absolute and relative weights at both colony and group of colonies scales. The value used to split mortality data in `bdy_processing_mortality` is the relative weight of mortality between the parcs for each group of colonies (`RW_group`).
```{r apportionning}
apportionning <- bdy_apportionning(max_foraging_range_km = foraging_ranges$max_km[which(foraging_ranges$species_latin==sp)],
                                   colonies = count_processed$colonies_sp,
                                   sea_area = sea_area,
                                   tbl_dist = distances
)

print(apportionning$RW_group)
```


This table can then be used in the [bdy_processing_mortality()] that distributes national mortality data among all colonies of a given species.
```{r apportion_mortality}
morta_iter_group <- bdy_processing_mortality(
      collision = FormattedMortality,
      season=seasons[which(seasons$Espèce==sp),][-1],
      n_iteration = 1000,
      RW_group = apportionning$RW_group
    )

```


## Model with no impact
In this step, we will create the demographic model that estimates species trends if there were no collisions with windfarms. Note that these trends are not expected to be very accurate on their own as the demographic models are quite simple, they are only meant to be used in comparison with the model with collisions.

```{r model_noimpact}
no_impact_output <- bdy_model_no_impact(count_data = count_processed$group_counts_sp,
                                        PI = count_processed$ppa_yr_sp,
                                        survival = vital_rates[vital_rates$species_latin == sp, "survival"],
                                        fecundity = vital_rates[vital_rates$species_latin == sp, "fecundity"],
                                        propRepro = vital_rates[vital_rates$species_latin == sp, "propRepro"],
                                        modelFile="C:/Users/TERRA/Documents/Projects/Birdynamic/app/0.Data/model_01.txt",
                                        nimble=T)
```

The output includes XX. 


## Model with impact
We will now create the model with impact from collisions with windfarms by applying the mortalities processed in the apportionning step to the demographic trends estimated with the model with no impact. 

```{r model_withimpact}
with_impact_output <- bdy_model_with_impact(n_group=nlevels(count_processed$colonies_sp$group),
                                            ny_data=ncol(count_processed$group_counts_sp),
                                            posterior=no_impact_output,
                                            mortality=morta_iter_group
)

```

The output includes XX. 

## Result table

```{r result_table}
Model_output <- list(
  SP = list(
    no_impact=with_impact_output$sc0,
    with_impact=with_impact_output$sc1,
    colonies=count_processed$colonies_sp,
    mortality=morta_iter_group,
    distance=round(distances)#,
    #distanceBIN=distanceBIN
  )
)
names(Model_output)[1] <- sp

# bdy_raw_res_tables(mod_out=Model_output)

```



# Grouped analysis for multiple species
The same analysis can be run using the grouped function `bdy_run_analysis`, which includes the steps described in the previous section, and allowing running the analysis for multiple species.

```{r analyses_multiple}
mod_outputs <- bdy_run_analysis(
  species = c("Morus bassanus", "Uria aalge", "Sterna dougallii"),
  countData = countData,
  foraging_ranges = foraging_ranges,
  mortality = FormattedMortality,
  parcs = parcs_L93
)

#mod_outputs
```

The output includes XX. 



# Display results
Whether models were run using the individual functions or the grouped function, the package includes functions to display the different results, following the outputs displayed in the Bird Dynamic Shiny app.

The tables returned from the function 





# Conclusion




