---
title: "Use BirdynamicPkg package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Use BirdynamicPkg package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```


# Introduction
The R package 'BirdynamicPkg' enables modelling the demographic impact of windfarms on seabird species following the methodology developed in the Bird Dynamic project. Check out the 'Read me' file to know more.


# Download and load the package

```{r setup, echo=F}
#devtools::install_github("BaptisteAverly/BirdynamicPkg", ref="develop")
library(BirdynamicPkg)

library(stochLAB)
library(rhandsontable)
library(popbio)
library(plyr)
library(tidyverse)
library(plotly)
library(rmarkdown)
library(markdown)
library(DT)
library(leaflet) # Needed for interactive map
library(leafem) # Needed for addMouseCoordinates
library(gdistance) # Needed to calculate shortest path distance
library(sf) # Needed for GIS work
library(raster) # Needed for GIS work (extent function)
library(units)
library(htmlwidgets) # Needed for onRender in leaflet map
library(reshape2) # Needed to prepare output tables
library(nimble)

#load("C:/Users/TERRA/Documents/Tempo_Bdy_loaded4vignette.RData")

# # TEMPO
# parcs_L93 <- st_read("C:/Users/TERRA/Documents/Projects/Birdynamic/Test_data_Shiny/Parcs_generic_names.shp") %>% st_transform(., st_crs(2154))
# parcs_L93$Facade <- st_join(st_point_on_surface(parcs_L93), LimFrance, join=st_intersects)$Facade
```


# Prepare data

## Species parameters
To run the analyses, parameters of the species of interest have to be included in the following tables. For each of these tables, the default values that we use in the Bird Dynamic Shiny app are included in the package, but you can decide to edit them with different values or adding species.

The table `vital_rates` includes the survival and reproduction parameters:
```{r data_vital}
vital_rates <- bdydata_vital_rates
head(vital_rates)
```


The table `foraging_ranges` includes the foraging range of each species:
```{r data_foraging}
foraging_ranges <- bdydata_foraging_ranges
head(foraging_ranges)
```


The table `seasons` includes the presence of each species at each month (see `help(bdydata_seasons)` to see details on attributes):
```{r data_seasons}
seasons <- bdydata_seasons
head(seasons)
```


## Mortality data
Mortality can be either calculated using the `stochLAB` package or directly loaded if they were calculated previously.

### Calculate mortality data
To create them with the `stoch_crm()` function from the `stochLAB` package, you must define the density of your species for each parc and month, and all the parc and seabird parameters. The values entered below are only for the example, do not consider them as verified default values. 

```{r calculate_mortality}

dens_dt <- data.frame(
  row.names = month.name,
  month=month.name,
  mean=c(5,5,5,10,10,10,10,10,10,5,5,5),
  sd=1
)

chord_profile <- data.frame(
  pp_radius=seq(0, 1, by=0.05),
  chord=c(0.69, 0.73, 0.79, 0.88, 0.96, 1, 0.98, 0.92, 0.85, 0.8, 0.75, 0.7, 0.64, 0.58, 0.52, 0.47, 0.41, 0.37, 0.3, 0.24, 0)
)

n_iterations <- 1000

morta_results <- stoch_crm(
  model_options = "1",
  n_iter = n_iterations,
  flt_speed_pars = data.frame("mean"=0.1, "sd"=0),
  body_lt_pars = data.frame("mean"=0.1, "sd"=0),
  wing_span_pars = data.frame("mean"=0.1, "sd"=0),
  avoid_bsc_pars = data.frame("mean"=0.99, "sd"=0),
  avoid_ext_pars = NULL,
  noct_act_pars = data.frame("mean"=0.2, "sd"=0),
  prop_crh_pars = data.frame("mean"=0,"sd"=0),
  bird_dens_opt = "tnorm",
  bird_dens_dt = dens_dt,
  flight_type = "flapping",
  prop_upwind = 0.5,
  n_blades = 3,
  air_gap_pars = data.frame("mean"=30,"sd"=0),
  rtr_radius_pars = data.frame("mean"=80, "sd"=0),
  bld_width_pars = data.frame("mean"=5.5,"sd"=0),
  bld_chord_prf = chord_profile,
  rtn_pitch_opt = "probDist",
  bld_pitch_pars = data.frame("mean"=2, "sd"=0.1),
  rtn_speed_pars = data.frame("mean"=10, "sd"=0.5),
  trb_wind_avbl = cbind(data.frame("month"=month.name),"pctg"=rep(95, 12)),
  trb_downtime_pars = cbind(data.frame("month"=month.name), "mean"=rep(6, 12), "sd"=rep(2, 12)),
  wf_n_trbs = 50,
  wf_width = 5,
  wf_latitude = 48.67,
  tidal_offset = 2,
  lrg_arr_corr = T,
  out_format = "draws",
  out_sampled_pars = FALSE,
  out_period = "months",
  season_specs = NULL,
  verbose = TRUE,
  log_file = NULL,
  seed = NULL
)

morta_format <- data.frame("espece_latin"="Larus fuscus",
                           "parc"="parc_1",
                           "mois"=rep(1:12, n_iterations),
                           "iteration"=rep(1:n_iterations, each=12),
                           "coefficient"=as.vector(t(morta_results[["opt1"]]))
                           )
```


### Load mortality data
Alternatively, mortality data that were compiled previously can be directly loaded, as long as they are formatted as in `bdydata_mortality_example` (which we will use in this vignette):

```{r load_mortality}
FormattedMortality <- bdy_check_mortality(bdydata_mortality_example, bdydata_mortality_example$espece_latin)$table
head(FormattedMortality)
```



## Parc data
Parc data must be provided to run the analyses, as a sf object (see `st_read(st_read)` for more details) using either points or polygons, projected in Lambert93 (EPSG:2154). As an example, we here use the `bdydata_parc_example` data included in the package. They must include a column indicating the seafront (for France, it has to be 'atlantic' or 'mediterranean'); this also has to be added as a column to the colony file.

```{r parc_example}
parcs_L93 <- bdydata_parcs_example %>% st_transform(., st_crs(2154))
parcs_L93$seafront <- "atlantic"
parcs_L93

colonies00_L93$seafront <- "atlantic"
```


ADD Check mortality names


# Analyse data

## Calculate distance between colonies and parcs
We calculate two types of distance: euclidian distance is used for species that cross land (e.g., gulls), shortest sea path is used for purely marine species (e.g., Northern Gannet). Both distances are calculated for all colonies (if at least one species is purely marine) with the [bdy_get_distances()] function, the type of distance to use then depends on species later in the analysis.

```{r calc_dist}
all_distances <- bdy_get_distances(colonies=colonies00_L93,
                                   parcs = parcs_L93,
                                   costMatrix=costMat,
                                   doShpa = T)

shortest_path_distances <- all_distances$shpa_dist
head(shortest_path_distances)

euclidian_distances <- all_distances$eucl_dist
head(euclidian_distances)

```


## Species analysis
The next steps of the analysis are applied species by species. In this vignette, we show a detailed example for one species, followed by how you can loop the analysis for multiple species.

For this example, we will select the Northern Gannett (**Morus bassanus**) and load the data saved during the previous step (TO ADAPT).
```{r select_species}
sp <- "Fou de Bassan"

load(paste0("C:/Users/TERRA/Documents/Projects/Birdynamic/app/0.Data/processed_data_per_species/clean_data_",sp,".RData"))
```


### Apportionning and process mortality data
The first step to apply is the apportionning, where mortalities due to collisions with wind turbines (modelled or loaded earlier in the pipeline) are attributed to the different colonies depending on their distance to the windfarms or the configuration of the coastline.

We first set the distance between parcs and colonies that include the selected species, keeping the shortest sea path if the species is purely marine, the euclidian distance otherwise.

```{r select_dist}
if(sp %in% moeurs_terrestres){
      selec <- which(row.names(euclidian_distances) %in% colonies$code_colonie)
      colo_parc_distances <- euclidian_distances[selec,,drop=F]
    }else{
      selec <- which(row.names(shortest_path_distances) %in% colonies$code_colonie)
      colo_parc_distances <- shortest_path_distances[selec,,drop=F]
    }

```


We then load sea areas (TO DETAIL).
```{r select_seaarea}
 if(sp == "Guifette noire"){
      sea_area <- rep(1, nrow(colonies))
      names(sea_area) <- row.names(colo_parc_distances)
    }else{
      load(paste0("C:/Users/TERRA/Documents/Projects/Birdynamic/app/0.Data/sea_area/sea_area_", sp, ".rda"))
    }

```


We can now apply the [bdy_apportionning()] function to run the apportionning calculations to set how mortality data should be distributed among colonies.
```{r apportionning}
apportionning <- bdy_apportionning(max_foraging_range_km = foraging_range_tbl$Max_km[which(foraging_range_tbl$Espece==sp)], #B
                                       colonies = colonies,
                                       sea_area=sea_area,
                                       tbl_dist=colo_parc_distances)

```


This table can then be used in the [bdy_processing_mortality()] that distributes national mortality data among all colonies of a given species.
```{r apportion_mortality}
morta_iter_group <- bdy_processing_mortality(
      collision = FormattedMortality,
      season=seasons[which(seasons$Espèce==sp),][-1],
      RW_group = apportionning$RW_group
    )

```


### Model with no impact
In this step, we will create the demographic model that estimates species trends if there were no collisions with windfarms. Note that these trends are not expected to be very accurate on their own as the demographic models are quite simple, they are only meant to be used in comparison with the model with collisions.

```{r model_noimpact}
no_impact_output <- bdy_model_no_impact(count_data=group_counts,
                                        PI=ppa_yr,
                                        survival=vital_rates[vital_rates$Espèce == sp, "Survie"],
                                        fecundity=vital_rates[vital_rates$Espèce == sp, "Fecund"],
                                        propRepro=prop_repro[vital_rates$Espèce == sp, "PropRepro"],
                                        modelFile="C:/Users/TERRA/Documents/Projects/Birdynamic/app/0.Data/model_01.txt",
                                        nimble=T)
```

The output includes XX. 


### Model with impact
We will now create the model with impact from collisions with windfarms by applying the mortalities processed in the apportionning step to the demographic trends estimated with the model with no impact. 

```{r model_withimpact}
with_impact_output <- bdy_model_with_impact(n_group=nlevels(colonies$group),
                                                ny_data=ncol(group_counts),
                                                posterior=no_impact_output,
                                                mortality=morta_iter_group)

```


### Use of grouped function and loop for multiple species
The same analysis can be run using the grouped function ADD_FUNCTION, which includes the steps described in the previous section (from selecting the correct distance table to the model with impact). This is convenient for instance if you need to loop the analyses across multiple species.

ADD EXAMPLE
```{r analyses_multiple}
with_impact_output <- bdy_model_with_impact(n_group=nlevels(colonies$group),
                                                ny_data=ncol(group_counts),
                                                posterior=no_impact_output,
                                                mortality=morta_iter_group)

```

## Display results





<!-- ## RAW -->

<!--     # Prepare binary distance > foraging -->
<!--     Foraging_range <- foraging_range_tbl$Max_km[which(foraging_range_tbl$Espece==sp)] -->
<!--     units(Foraging_range)<-"km" -->
<!--     distanceBIN <- colo_parc_distances<Foraging_range ; rownames(distanceBIN)<-rownames(colo_parc_distances) ;  names(distanceBIN) <- names(colo_parc_distances) -->
<!--     model_output[[sp]] <- list(no_impact=with_impact_output$sc0, with_impact=with_impact_output$sc1, colonies=colonies, mortality=morta_iter_group, distance=round(colo_parc_distances), distanceBIN=distanceBIN) -->

<!--     rm(colonies,group_counts,ppa_yr) -->
<!--   } #end species loop -->



<!-- # Display results -->

<!-- ## RAW -->

<!--   ### Prepare tables (Subpop and National) -->
<!--   Raw_ResTables <- bdy_raw_res_tables(model_output) -->
<!--   Stored_results$Res_list$Raw_ResTables <- Raw_ResTables -->

<!--   Pretty_table_national <- bdy_pretty_result_table(Raw_ResTables, type="national") -->
<!--   Stored_results$Res_list$DT_table_national <- Bdy_InteractiveResultTable(Pretty_table_national, type="national") -->

<!--   ### Prepare summary figures -->
<!--   # Trends -->
<!--   Summary_Figures<- bdy_summary_figure(Raw_ResTables, Pretty_table_national) -->
<!--   Stored_results$Res_list$Summary_Gyeartrends <- Summary_Figures -->
<!--   Height_plot <- length(unique(ggplot_build(Summary_Figures)$layout$layout$ROW)) * 200 + 150 -->
<!--   Stored_results$Res_list$Summary_Gyeartrends_interactive <- ggplotly(Summary_Figures, tooltip="text", height=Height_plot, dynamicTicks = FALSE) %>% -->
<!--     plotly::layout(xaxis = list(fixedrange = TRUE), xaxis2=list(fixedrange = TRUE), xaxis3=list(fixedrange = TRUE), xaxis4=list(fixedrange = TRUE), xaxis5=list(fixedrange = TRUE), yaxis = list(fixedrange = TRUE)) %>% -->
<!--     config(displayModeBar = F) -->

<!--   # Interactive map -->
<!--   Stored_results$Res_list$Summary_Map <- Bdy_SummaryMap(model_output, Raw_ResTables, parcs_L93, BufferForaging) -->

<!--   # Mortalite par parc -->
<!--   Stored_results$Res_list$MortaParc <- Bdy_MortaParc(Raw_ResTables, FormattedMortality()[(FormattedMortality()$espece %in% sel_sp)  & (FormattedMortality()$parc %in% input$selected_parcs),]) -->



# Conclusion




